\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{setspace}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{float}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{enumitem}
\usepackage{caption}

\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    urlcolor=blue,
    citecolor=blue
}

\title{Project 4 Final Report\\End-to-End Machine Learning on an Open E-Commerce Dataset}
\author{
    Team Member 1 \and
    Team Member 2 \and
    Team Member 3
}
\date{\today}

\begin{document}

\maketitle
\onehalfspacing

\begin{abstract}
This report presents an end-to-end machine learning study built on an open e-commerce dataset. The project is designed to mirror a practical data science workflow, beginning with data acquisition and cleaning, continuing through exploratory analysis and feature engineering, and ending with supervised model development, comparison, and selection. In addition to predictive modeling, the analysis incorporates unsupervised learning to better understand latent structure in the data and to support downstream modeling decisions. The goal of the project is not only to obtain strong predictive performance, but also to justify modeling choices, assess tradeoffs among alternative approaches, and communicate the results in a clear and reproducible way.
\end{abstract}

\section{Introduction}

E-commerce platforms generate rich behavioral and transactional data that can support a wide range of business decisions, including customer targeting, operational planning, product strategy, and service quality monitoring. Because these systems capture interactions among customers, products, payments, logistics, and outcomes, they provide a useful setting for studying the full machine learning pipeline from raw data to actionable conclusions.

In this project, we analyze an open e-commerce dataset to investigate a predictive question of practical interest. Our workflow emphasizes three ideas. First, real-world data rarely arrive in a form that is immediately ready for modeling, so data cleaning and preprocessing are essential parts of the analysis rather than minor preliminary steps. Second, exploratory analysis and unsupervised learning can reveal structure that would be difficult to detect from raw tables alone. Third, model selection should reflect not only predictive performance but also interpretability, robustness, and fit to the project context.

To address these goals, we organize the report around the major stages of an end-to-end data science workflow. We begin by describing the data source and the preparation steps needed to transform the raw data into an analysis-ready form. We then use visualization and unsupervised learning to better understand the data and motivate feature engineering choices. Next, we train and compare multiple supervised learning models using appropriate validation procedures and evaluation metrics. Finally, we select a preferred model, interpret the results, and discuss the limitations and possible extensions of the analysis.

\section{Project Motivation and Predictive Question}

\textit{Placeholder: State the business or analytical motivation, define the target variable, and explain why the prediction task matters.}

\section{Data Acquisition and Preparation}

This project uses the Brazilian E-Commerce Public Dataset by Olist, an open multi-table dataset made available through Kaggle. The dataset contains information on customers, orders, order items, products, sellers, payments, geolocation, and customer reviews. Its relational structure makes it well suited for an end-to-end machine learning project because the raw data must be merged, cleaned, and transformed before analysis-ready features can be constructed.

For this project, the raw tables will be downloaded from the public Olist release and stored in a reproducible folder structure. The preparation stage will focus on identifying primary and foreign-key relationships across tables, standardizing date and time fields, checking for duplicate records, and resolving missing or inconsistent values. Because the dataset combines transactional, geographic, logistical, and review information, special attention will be paid to join logic and leakage prevention so that only information available at prediction time is used in supervised models.

\textit{Placeholder: Add the exact download date, licensing note, table list, row counts, and the final cleaning decisions after the pipeline is implemented. Include missing-value handling, type conversion, duplicate handling, outlier rules, and any filtering.}

\section{Exploratory Data Analysis}

\textit{Placeholder: Summarize key distributions, relationships, anomalies, and initial business insights. Add figures and descriptive statistics here.}

\section{Unsupervised Learning}

\textit{Placeholder: Describe clustering, PCA, UMAP, or another unsupervised method used to reveal structure or support feature engineering.}

\section{Feature Engineering and Preprocessing}

\textit{Placeholder: Document feature construction, encoding, scaling, dimensionality reduction, train-validation-test design, and leakage prevention.}

\section{Supervised Models}

\subsection{Logistic Regression}
\textit{Placeholder: Describe the logistic regression model, regularization choices, tuning strategy, and evaluation results.}

\subsection{Gradient-Boosted Trees}
\textit{Placeholder: Describe the gradient-boosted tree model, implementation choice, tuning strategy, and evaluation results.}

\subsection{Feed-Forward Neural Network}
\textit{Placeholder: Describe the neural network architecture, training procedure, tuning strategy, and evaluation results.}

\section{Model Comparison and Final Selection}

\textit{Placeholder: Compare models using performance, interpretability, robustness, and practical usefulness. Explain why the final model was chosen.}

\section{Discussion}

\textit{Placeholder: Interpret the main findings, discuss limitations, and outline possible future improvements.}

\section{Reproducibility}

\textit{Placeholder: Provide enough detail for another reader to reproduce the analysis, including software, package versions, folder structure, and execution steps.}

\section{Team Contributions}

\begin{itemize}[leftmargin=1.5em]
    \item Team Member 1: \textit{Placeholder}
    \item Team Member 2: \textit{Placeholder}
    \item Team Member 3: \textit{Placeholder}
\end{itemize}

\section{Conclusion}

\textit{Placeholder: Summarize the project question, key findings, and the final takeaway in a short closing section.}

\begin{thebibliography}{9}

\bibitem{dataset}
Olist. \textit{Brazilian E-Commerce Public Dataset}. Kaggle. \url{https://www.kaggle.com/datasets/olistbr/brazilian-ecommerce}

\bibitem{method}
\textit{Placeholder: Add methodological references here as needed.}

\end{thebibliography}

\appendix

\section{Additional Tables and Figures}

\textit{Placeholder: Move large tables, supplementary plots, or extended results here.}

\end{document}
